---
title: "Global ETF Map Presentation"
output:
  html_notebook: default
---

## How we'll get there

- R Notebook for data import, wrangling, and test visualizations
- .RDat File
- Flexdashboard to build our interactive Shiny app

```{r, message = FALSE}
# Let's build a dataframe to store these ticker symbols, country names and YTD numbers.
library(tidyverse)
library(tidyquant)

tickers <-  c("ECH", "EDEN", "EFNL", "EIDO", "EIRL", "EIS", "ENOR", "ENZL", 
"EPHE", "EPOL", "EPU", "ERUS", "EWA", "EWC", "EWD", "EWG", "EWH", 
"EWI", "EWJ", "EWK", "EWL", "EWM", "EWN", "EWO", "EWP", "EWQ", 
"EWS", "EWT", "EWU", "EWW", "EWY", "EWZ", "EZA", "FXI", "ICOL", 
"INDA", "KSA", "QAT", "THD", "TUR", "UAE")

name <-   c("Chile", "Denmark", "Finland", "Indonesia", "Ireland", "Israel", 
"Norway", "New Zealand", "Philippines", "Poland", "Peru", "Russia", 
"Australia", "Canada", "Sweden", "Germany", "Hong Kong", "Italy", 
"Japan", "Belgium", "Switzerland", "Malaysia", "Netherlands", 
"Austria", "Spain", "France", "Singapore", "Taiwan", "United Kingdom", 
"Mexico", "Korea", "Brazil", "South Africa", "China", "Colombia", 
"India", "Saudi Arabia", "Qatar", "Thailand", "Turkey", "United Arab Emirates"
)

etf_ticker_country <- data_frame(tickers, name)

etf_ticker_country
```

 

```{r, warning = FALSE, message = FALSE, eval = FALSE}
# getSymbols is part of the 'quantmod' package.

# Using getSymbols to import the ETF price histories will take a minute or two or 
# five - 41 time series is a lot of data. 

# Let select just the Adjusted prices of the ETFs and merge them into a list.
# We'll use map and the piper operator for that purpose. Again, this is for testing. It's not 
# going into production in our app.

etf_prices <- 
  getSymbols(etf_ticker_country$tickers, auto.assign = TRUE, warnings = FALSE) %>%
  # Let's use the map function to get just the Adjusted prices
  map(~Ad(get(.))) %>% 
  # Now use the reduce() function to combine to one xts object
  reduce(merge) %>% 
  # some name cleanup
  `colnames<-`(etf_ticker_country$name)

# Take a peek at the last 5 rows of each of the time series, 
# just to make sure it looks complete.
library(timetk)
etf_prices_no_na <- 
  etf_prices %>%
  tk_tbl(preserve_index = TRUE, rename_index = "date") %>% 
  na.omit() %>% 
  filter(date > "2016-12-31") %>% 
  gather(name, price, -date) %>% 
  group_by(name) %>% 
  summarise(ytd = round(last(price) - first(price)/last(price), 4))

etf_ticker_country <- merge(etf_prices_no_na, etf_ticker_country, by = "name", all.x = TRUE)

tail(etf_prices, n = 5)
```



```{r, message = FALSE, warning = FALSE}
library(rnaturalearth)
library(sf)
world <- ne_countries(type = "countries", returnclass = 'sf')
```



```{r, message = FALSE}
# For the sake of brevity, let's look at just the first 6 countries, 
# their GDP estimates and the stage of their economy.
head(world[c("name", "gdp_md_est", "economy")], n = 6)

#If you want to peek at the longitude and latitude, use world$geometry
```



```{r, warning = FALSE, message = FALSE}
library(leaflet)  

# Create a palette with different shades of blue for different
# GDP estimates.

gdpPal <- colorQuantile("Blues", world$gdp_md_est, n = 20)
```




```{r, warning = FALSE, message = FALSE}
# Make a popup object.
# Notice we're referencing the column names with the '$', same as we would with a non-spatial dataframe.
economyPopup <- paste0("<strong>Country: </strong>", 
                world$name, 
                "<br><strong>Market Stage: </strong>", 
                 world$economy)
```



```{r, warning = FALSE, message = FALSE}
# Build our leaflet map object.

leaf_world_economy <- leaflet(world) %>%
  addProviderTiles("CartoDB.Positron") %>% 
  setView(lng =  20, lat =  15, zoom = 2) %>%
      addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = .7, color =
      ~gdpPal(gdp_md_est), layerId = ~name, popup = economyPopup)

# Display that object below.

leaf_world_economy
```



```{r, warning = FALSE, message = FALSE}
library(sf)
# Once we run this line of code, our ticker symbols and ytd numbers will be added
# to the simple features dataframe.

world_etf <- merge(world, etf_ticker_country, by = "name", all.x = TRUE)

head(world_etf)
```



```{r, warning = FALSE, message = FALSE}
# Create a palette with different shades of red yellow green for different
# year-to-date performances. Previously, we shaded by 'world$gdp_md_est', now
# we'll shade by 'world_etf$ytd'.

ytdPal <- colorQuantile("RdYlGn", world_etf$ytd, n = 20)
```


The new shading is nice, but let’s also have the popup display the exact year-to-date performance percentage for any detail-oriented users.

```{r, warning = FALSE, message = FALSE}
# Create a popup that displays the year-to-date performance.

ytdPopup <- paste0("<strong>Country: </strong>", 
                world_etf$name,
                "<br><strong> Year-to-date: </strong>", 
                world_etf$ytd, "%")
```
 


```{r}

leaf_world_etf <- leaflet(world_etf) %>%
  addProviderTiles("CartoDB.Positron") %>% 
  setView(lng =  20, lat =  15, zoom = 2) %>%
      addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = .7,
                  
      # The next line of code is really important for creating the map we want to use later.       
      
      color =~ytdPal(ytd), layerId = ~tickers, popup = ytdPopup)

leaf_world_etf
```


This map is going to play a crucial role in our Shiny app and the last step is to save the leaflet object for use in our flexdashboard. 


```{r}
# Save the shapefile but NOT the ETF price data to a .RDat file.
# We'll load this file and use it in our Shiny app.

save(leaf_world_etf, file = 'etfData.RDat')
```



